Bayesian non-parametric inference for multivariate peaks-over-threshold models

### 2023-05-17 pointbypoint.tex

\newcommand{\comment}[1]{\textcolor{teal}{#1}}
\newcommand{\bruno}[1]{\textcolor{red}{#1}}

\begin{document}

\section*{Review 1}
\begin{itemize}
\item \emph{Moment constraint}: As can be seen, for example, in \cite[Equation 4.1]{einmahl2009}
    there is a moment constraint on the angular measure for any $\mathcal{L}_p$ norm with 
    $p \in [1, \infty]$ (nb including $p = \infty$).  Hence, why is there no moment constraint in 
    the submitted 
    manuscript? Some papers on inference for the angular measure have not taken into consideration 
    that constraint (e.g. \cite{einmahl2001}), though most modern approaches do account for it. If 
    the moment constraint is being ignored here, I would be willing to overlook this provided that 
    this is openly acknowledged---as well as what are the consequences of this. Otherwise, this needs 
    to be further clarified.

    % \comment{Both reviewers comment on the moment constraint.  I forget if it's this one, or
    % the reference provided by the other reviewer, but I don't think they understood the reference
    % directly; they were using the $\mathcal{L}_1$ norm in $p$ dimensions, rather than the 
    % $\mathcal{L}_p$ norm as described in our paper.  A massive difference in meaning of notation.
    % I'll find the reference again and confirm.}

@@

We thank the referee for calling our attention to
    the issue of moment restrictions. The approach considered in the
    paper follows that of \cite{rootzen2018}, consisting on modeling the
    distribution of the extreme observations, conditional on them being
    above a threshold, say $H$.  The approach considered in
    \cite{EiSe2009} focuses in the limiting joint exponent measure. More
    precisely, for the two dimensional case, they assume that
    \[
        sPr(s^{-1}(Z_1,Z_2) \in A) \rightarrow \mu(A),
    \]
    where $\mu$ is the exponent measure. The moment conditions are
    the result of assuming that both marginals of $\mu$ correspond to a
    standard Pareto distribution. Such assumption is not compatible
    with the conditional distribution approach considered in our paper.
    In fact, as noted in \cite{KiRoSeWa2019}, typically the margins of
    $H$ are not univariate Pareto, due to the difference between the 
    conditioning events in the joint and the marginal cases. Thus, no
    moment conditions arise in our model. This has now been clarified in the paper.

We note in passing that the moment conditions for the spectral distribution, say 
    $\Phi$, in the two-dimensional case, using the infinity norm, imply that 
    \[
        \int_0^{\pi/4} \Phi(d\theta) = \int_{\pi/4}^{\pi/2}
    \Phi(d\theta).
    \]
    The former looks like a fairly strong symmetry restriction that is unlike to 
    be realistic in practice. Thus, imposing moment conditions would probably 
    result in a very restrictive model.

\item \emph{Main contribution}: While I appreciate the interesting construction 
    based on the DPM of projected gammas (which itself could be of use for other 
    contexts beyond extremes?), one wonders about what would be the main added 
    value of it with respect to other existing Bayesian approaches 
    (\cite{boldi2007},\cite{guillotte2011},\cite{SaNa2014},\cite{hanson2017}).
    Keeping in mind the computational scope of Statistics and Computing, one 
    wonders if the main advantage is, for example, a computational one? This 
    needs to be clarified—especially in light of the previous comment on the 
    moment constraint. Indeed, one could use any of the previous approaches with 
    pseudo-angles based on the infinity norm.

We have clarified the distinction between our approach and the prior contributions
    mentioned.

    % \comment{Hanson is clearly different.  They're using Bernstein polynomials 
    % effectively defined on $\mathbb{S}_1^{d-1}$.  We can demonstrate that 
    % projected gammas built on $\mathbb{S}_{p}$, $p \geq 10$ and $\mathbb{S}_1$ 
    % create \emph{very} different gamma shape parameters, for data created by 
    % the \emph{same} distribution.  I need to check again the other references, 
    % but based on their follow-on comment, I'm guessing the previous point needs 
    % to be addressed before we can properly justify this point.}

    % \bruno{Based on our conversation, you need to look carefully into each one 
    % of these references. it seems that most of them do not develop a PoT 
    % approach. But also, we use a kernel whose natural support is the positive 
    % cone (without using transformations) to build a mixture WRT to a random
    % measure. This provides the flexibility of mixture for a methods that has 
    % computational advantages and scales to moderately large problems, as 
    % illustrated in the paper.}

    \comment{\cite{Hanson2017}: Bernstein polynomials on $\mathbb{S}_1^{d-1}$, 
    modelling multivariate Frechets.  \cite{boldi2007}: Marginal Frechets, also 
    operating on $\mathbb{S}_1^{d-1}$.  \cite{guillotte2011}: A strictly bivariate model.  
    \cite{SaNa2014}: }

@@

\item \emph{Numerical accuracy}:  Keeping in mind the computational scope of 
    Statistics and Computing one would hope for more numerical work than what is 
    presented in §5.1. As far as I understand what is presented are single run 
    experiments for each $d$? Monte Carlo evidence on the performance of the 
    procedure would be appreciated. On top of this, the current numerical exercise
    ignores the uncertainty that stems from step 1 of the procedure 
    (i.e. estimating the margins).

To address this concern, we have revised the simulation study to feature the 
    average rise in energy score over baseline, averaged over 10 replications.  
    The simulated datasets are generated as mixtures of multivariate lognormals, 
    projected onto the unit sphere.

    % \comment{Regarding the uncertainty from step 1, I think we agreed that was 
    % basically intractible. We may need to explicitly specify as such, and that a 
    % practicioner must accept the numerical uncertainty that arises from step 1.  
    % That said, I seem to recall the numbers observed being relatively robust to 
    % choice of threshold (for large threshold), the difference mainly coming down 
    % to computation time.}

    % \bruno{I think we need to conduct a more comprehensive suite of experiments. 
    % Regarding the uncertainty coming from Step1, we just need to acknowledge that 
    % we are ignoring it.}

@@

        \item Another issue is that it is not clear in the present form of the paper, what is the
        added value of fitting this parametric model with the Bayesian framework and the Gibbs 
        sampler, if the output is mainly bivariate densities and extremal dependence coefficients 
        which could be obtained by standard non parametric methods with very low computational cost?

        \bruno{This is an unfair comment. We do use the full multivariate distributiont to come up
        with the results we present. I'll work on an answer to this one.}
    \end{enumerate}

\item \emph{Comparison with other methods}: The only comparison study displayed in the paper 
    takes place within the suggested model, only with different prior choice. This is not sufficient 
    for performance comparison in a methodological paper. One may e.g. think of comparing with the
    results obtained with the empirical angular measure or smoothed versions of it in terms of 
    extremal dependence and conditional survival probability.

    We have included for comparison some other models established on $\mathcal{S}_{1}^{d-1}$;
        notably the pairwise betas model developed in \cite{sabourin2013}, as well
        as the dirichlet distribution.  We have investigated several packages to

### 2023-05-17 introduction.tex
OL 2023_10_15
minor update

During the last decade or so, much work has been done in the exploration of the definition 
  and properties of an appropriate generalization of the univariate GP distribution 
  to a multivariate setting.  To mention some of the papers on the topic, the work of
  \citep{rootzen2006} defines the generalized Pareto distribution, with further analysis on these classes
  of distributions presented in \cite{falk2008} and \cite{michel2008}.  A recent review of the state
  of the art in multivariate peaks over threshold modelling using generalized Pareto is provided in
  \cite{rootzen2018} while \cite{RoSeWa2018a} provides insight on the theoretical properties of possible parametrizations. These are use in \cite{KiRoSeWa2019} for likelihood-based models for PoT estimation.
  A frequently used method for describing the dependence
  in multivariate distributions is to use a copula. \cite{renard2007}, and \cite{falk2019} provide successful examples of this approach in an EVT framework. \cite{ferreira2014} presents a constructive definition of the Pareto process, that generalizes the GP to an infinite dimensional setting. It consists of decomposing the process into independent radial and angular components. Such an approach can be used in the finite dimensional case, where the angular component 
  contains the information pertaining to the dependence structure of the random vector.
  Based on this definition, we present a novel approach for modelling  the angular component
  with families of distributions that provide flexibility and can be applied in a moderately large dimensional setting.
  Our focus on the  angular measure is similar to that in \cite{SaNa2014} and \cite{HaCaCh2017}, that consider Bayesian non-parametric approaches. Yet, our approach avoids the problem of dealing with the so called moment constraint by considering a constructive definition of the multivariate GP based on the infinity norm.

@@

\section{A multivariate PoT model\label{sec:multivariatepot}}
%To develop a multivariate PoT model for extreme values consider a 
%$d$-dimensional random vector $\bm{W} = (W_1, \ldots ,W_d)$  with 
%marginal distributions $F_1(w_1), \dots , F_d(w_d)$, and define its
%standardization as $\bm{Z} = (Z_1, \ldots , Z_d)$, where $Z_i = 1/ (1 - F_i(W_i)), \; i=1, \ldots , d$.
%Following the discussion in \cite{goix2017} we assume a form of 
%regular variation, consisting on assuming the existence of a 
%limiting measure $\mu$ on $(0,\infty]^d$, such that.
%\begin{equation}\label{eq:reg-var}
%        \lim\limits_{n\to\infty}n\prob{\frac{1}{n}\bm{Z}\not\leq \bm{z}} = \mu\left([\bm{0},\bm{z}]^c\right),
%\end{equation}
%where the inequality and the interval are taken componentwise. 
%In this setting, $\mu$ is the asymptotic distribution of $\bm{Z}$ 
%in extreme regions, and it is  referred to as the \emph{exponent
%measure}. In addition, we assume that $\lim_{n\rightarrow\infty} 
%n\prob{\max\{Z_1, \ldots , Z_d\}>n} = 1$. Equation \ref{eq:reg-var}
%implies that $\mu$ has the homogeneity property $\mu(t\cdot) = 
%\frac{1}{t}\mu(\cdot)$i, and it can be shown that  $\mu$ 
%can be decomposed into a radial and an angular component. In fact, 
%for $\bm{z} = (z_1, \ldots , z_n) \in (0,\infty]^d$, let $r(\bm{z}) = 
%||\bm{z}||_\infty$ and $\bm{v}(\bm{z}) = \bm{z}/r(\bm{z})$. Define 
%$\Phi(B) = \mu(\{\bm{z} : r(\bm{z})>1, \bm{v}(\bm{z}) \in B\})$, for any
%$B\subset \mathbb{S}_{\infty}^{d-1}$, the positive orthant of the unit 
%sphere in infinite norm. $\Phi$ is referred to as the {\em angular 
%measure}. Then 
%\begin{equation}\label{eq:exp-measure}
%    \mu\left(\left\lbrace  \bm{z} \in [0,\infty)^d : \lVert \bm{z} \rVert_{\infty} > r,\;\bm{z}/\lVert\bm{z}\rVert_{\infty} \in B\right\rbrace \right) = r^{-1}\Phi(B).
%\end{equation}
%Thus, the radial and angular components of the exponent measure are
%independent. To obtain a multivariate peak over threshold model we 
%consider the radial and angular component of the random variable $\bm{Z}$
%denoted, respectively, as $R = \max_{i} Z_i$ and $\bm{V} = \bm{Z}/R$. 
%Then, using Equations \ref{eq:reg-var} and \ref{eq:exp-measure}, we have
%that

%To obtain a statistical approach to estimating $\mu$, we follow the approach proposed in .
%Define  , and assume that it is distributed as a standard Pareto, and thus $R>1$ with probability one. Also, define , and assume that $\bm{V}$ has probability distribution $\Phi$ concentrated in $\mathbb{S}_{\infty}^{d-1}$, the positive orthant of the unit sphere in infinite norm. Also, assume that $E(V_i)>0$, and that $ \bm{Z}$ and $R$ are independent. 
%We refer to $R$ and $\bm{V}$ as, respectively, the  of $\bm{Z}$. Under those assumptions,  the convergence stated in Equation \ref{eq:reg-var} holds with limit measure $\mu$ determined by

%\citep[e.g.][applied to the finite-dimensional case]{ferreira2014}
%\[
%    \mu\left(\left\lbrace  \bm{z} \in [0,\infty)^d : \lVert \bm{z} \rVert_{\infty} > r,\;\bm{z}/\lVert\bm{z}\rVert_{\infty} \in A\right\rbrace \right) = r^{-1}\Phi(A)
%\]
%for $r > 0$ and for Borel sets $A\subseteq \mathbb{S}_{\infty}^{d-1}$.

%    $\mu(tA) = \frac{1}{t}\mu(A)$.  Define  $R = \max_{i} Z_i$, and assume that $R>1$ with probability one. 
%    Also, define $V_i = Z_i/\max_i Z_i$, and assume that $E(V_i)>0$. These two assumptions together with the 
%    homogeneity of $\mu$ correspond to the second condition in Theorem 2.1 of
%    \cite{ferreira2014}, applied to the finite-dimensional case.
%    Thus, we have the factorization $\bm{Z} =  R\bm{V}$, where $R$ and $\bm{V}$  are independent. Moreover,
%    for some measure $\Phi$, $\mu(\cdot)$ is factorized as
%    \temp{\begin{equation}\label{mufact}
%        %\mu\left( \left\lbrace\bm{Z} : V \in A , R > r\right\rbrace \right) = r^{-1}\Phi(A).
%        \mu\left(\left\lbrace \bm{z}\in \mathbb{R}_+^d : R(\bm{z}) > r, V(\bm{z}) \in A\right\rbrace\right) = r^{-1}\Phi(A)
%    \end{equation}
%    for $R(\bm{Z}) > 0$, and Borel sets $A \subseteq \mathbb{S}_{\infty}^{d-1}$.  Hereafter $R$ is referred
%    to as the radial component, and $\bm{V}$ the angular component of $\bm{Z}$
%    \st{$R$ describes the radial component, and $\bm{V} \in {\mathbb S}_{\infty}^{d-1}$ the 
%    angular component of $\bm{Z}$.}}  Here ${\mathbb S}_\infty^{d-1}$ denotes the positive
%    orthant of the unit sphere in infinite norm.  $\Phi(\cdot)$ is referred to as the \emph{angular measure}.   
%    From Equation (\ref{mufact}) we see that $R$ is distributed as a standard Pareto random variable. Conditioning on $R$, we have that
%    \begin{equation} \label{MGP}
%        \prob{\bm{V} \in B \mid R > r} = \frac{r\prob{\bm{V} \in B, R > r}}{r\prob{R > r}}\hspace{0.2cm}
%            \xrightarrow[r\to\infty]{~} \hspace{0.2cm} \Phi(B),
%    \end{equation}
%as $\lim_{r\rightarrow\infty} r\prob{R>r} = 1$.
%    Thus, the angular measure corresponds to the distribution of $\bm{V}$, conditional on $R$ being large.
%    As such, it controls the dependence structure of $\bm{Z}$ in the tails. 
%    In fact, assuming that $E(V_i)>0$ 
%    and following \cite{ferreira2014}, in the limit the distribution function of $\bm{Z}$ is given as
%    \begin{equation*}
%        \prob{Z_1\leq z_1, \ldots ,Z_d\leq z_d} = \expect{\bigvee_{i=1}^d
%    \frac{V_i}{z_i\wedge 1}} - \expect{\bigvee_{i=1}^d
%    \frac{V_i}{z_i}}
%    \end{equation*}
%    where the expectation is taken with respect to $\Phi$
%In view of this, to obtain a PoT model we consider a limiting 
%distribution for $\bm{Z}$ that corresponds to the distribution
%of the product of two independent random variables $\tilde{R}$ 
%and $\tilde{\bm{V}}$, with distributions, respectively, standard 
%Pareto and $\Phi$. \cite{ferreira2014} refer to such model as a 
%simple  Pareto process indexed on a finite dimensional set.  We 
%seek a flexible model for the distribution
%of $\tilde{\bm{V}} \in {\mathbb S}_{\infty}^{d-1}$.

%An alternative approach to justify the above mentioned limiting
%distribution, that provides a link to the multivariate GEV and the
%theory of max-stable distributions, is discussed 
%in Remark 1 of \cite{RoSeWa2018a}. More specifically, 
%using the notation in Theorem 6 of \cite{RoSeWa2018a}, define $\bm{S} 
%=  \log(\tilde{\bm{V}})$.  Then, according to Definition 5, $\bm{S}$ 
%is a spectral  random vector. Taking  $E \sim \text{Exp}(1)$, 
%independent of  $\tilde{\bm{V}}$,  we have that $\bm{Y} = \bm{S} + E$  
%follows a multivariate GP, according  to Theorem 6.  Then, using 
%Proposition 4,  $\tilde{\bm{V}}\tilde{R} = e^{\bm{Y}}$ will also follow
%a multivariate GP. Notice that $\tilde{R} = e^E$ is distributed as 
%a standard Pareto random variable. This is one of several stochastic
%representations of a multivariate GP presented in \cite{RoSeWa2018a}.

To develop a multivariate PoT model for extreme values consider a 
$d$-dimensional random vector $\bm{W} = (W_1, \ldots ,W_d)$ with
cumulative distribution $F$. Following \cite{RoSeWa2018a}, assume 
that there exists sequences of vectors $\bm{a}_n$ and $\bm{b}_n$,
and a $d$-variate distribution $G$ such that 
$\lim_{n\rightarrow\infty} F^n(\bm{a}_n \bm{w} + \bm{b}_n) = 
G(\bm{w})$. $G$ is a $d$-variate generalized extreme value 
distribution. It follows that
\begin{equation*}
    \begin{aligned}
    \lim\limits_{n\rightarrow\infty} &\prob{\bm{a}^{-1}_n (\bm{W} - \bm{b}_n) 
   \leq \bm{w} \mid \bm{W} \not\leq \bm{b}_n)} \\ 
     &\hspace{0.5cm}= \frac{\log G(\bm{w} \wedge \bm{0}) 
    - \log G(\bm{w}) }{\log G(\bm{0})} = H(\bm{w}).
    \end{aligned}
\end{equation*}
$H$ a multivariate Pareto distribution, and corresponds to the joint
distribution conditional on exceeding a multivariate threshold. In addition to
a copula function, $H$ depends on two $d$-dimensional vectors of 
parameters that we denote as $\bm{\xi}$ for the shapes and $\bm{\sigma}$ 
for the scales. \cite{RoSeWa2018a} provide a number of stochastic 
representations for $H$.  In particular, denote as $Z$ a
random variable with distribution $H$ where $\bm{\xi}= 1$ and 
$\bm{\sigma} = 0$.  Then, Remark 1 justifies the representation in 
\cite{ferreira2014}, giving $\bm{Z} = \bm{V}R$
where $R$ and $\bm{V}$ are independent. $R = \|\bm{Z}\|_\infty$ is 
distributed as a standard Pareto random variable, and $\bm{V} = 
\bm{Z}/\|\bm{Z}\|_\infty$ is a random vector in  
$\mathbb{S}_{\infty}^{d-1}$, the positive orthant of the 
unit sphere under $\mathcal{L}_{\infty}$ norm, with distribution $\Phi$. 
This representation is central to the methods proposed in this paper.
$R$ and $\bm{V}$ are referred to, respectively, as the {\em radial} 
and {\em angular} components of $H$. The angular measure controls 
the dependence  structure of $\bm{Z}$  in the tails. In view of 
this, to obtain a  PoT model we seek a flexible model for the 
distribution of $\bm{V} \in {\mathbb S}_{\infty}^{d-1}$.

@@

%\subection{Pairwise Extremal Dependence\label{subsec:ped}}
%%A measure that is used to characterize the strength of the dependence,
%%in the tail, for two random variables $Z_1$ and $Z_2$, with marginal
%%distributions $F_1$ and $F_2$ is given by \citep{coles2001}.
%%\[
%%    \chi_{12} = \lim\limits_{u\uparrow 1} \prob{F_1(Z_1)>u\mid F_2(Z_2)>u}  
%%\]
%%$\chi_{12}$ provides information about the distribution of extremes for the variable $Z_1$
%%given that $Z_2$ is very large.  When $\chi_{12}>0$, $Z_1$ and $Z_2$ 
%%are said to
%%be asymptotically dependent, otherwise they are asymptotically
%%independent. The following result provides the asymptotic dependence coefficient between two components of $\bm{Z}$
%%after PoT limit. 
%%\begin{prop}\label{ppchi}
%%Suppose that $\bm{Z} = R\bm{V}$ with $R\sim Pa(1)$,
%%$\prob{V_\ell > 0} = 1$ and $\expect{V_\ell}$ exists, for $\ell=1, \ldots ,d$, then
%%\begin{equation}
%%    \label{eqn:chi_ij}
%%	\chi_{\jmath\ell} = \expect{\frac{V_\jmath}{\expect{V_\jmath}} \wedge \frac{V_\ell}{\expect{V_\ell}}}
%%\end{equation}
%%\end{prop}
%%{\em Proof:}
%%Denote as $F_\ell$ the marginal distribution of $Z_\ell$. To obtain $\chi_{\jmath\ell}$ we need $Pr(Z_\jmath>z_\jmath,Z_\ell>z_\ell)$, where $z_\ell =
%%F_\ell^{-1}(u) = \expect{V_\ell}/(1 - u), \;\ell=1,\dots, d$.
%%Using the fact that
%%$V_\ell>0, \forall \ell$ almost surely, we have that the former is equal to
%%\begin{equation*}
%%    \begin{aligned}
%%    &\hspace{-0.5cm}\prob{R>\frac{z_\jmath}{V_\jmath}\vee\frac{z_\ell}{V_\ell}}
%%    = \expect{1\wedge\left(\frac{z_\jmath}{V_\jmath}\vee\frac{z_\ell}{V_\ell}\right)^{-1}}\\
%%    &= \expect{\frac{V_\jmath}{z_\jmath}\wedge\frac{V_\ell}{z_\ell}}
%%    = (1 - u)\expect{\frac{V_\jmath}{\expect{V_\jmath}}\wedge\frac{V_\ell}{\expect{V_\ell}}},
%%    \end{aligned}
%%\end{equation*}
%%where the second identity is justified by the fact that $V_i$ is
%%bounded and $z_i\rightarrow\infty$. The proof is completed by noting
%%that $\prob{F_i(Z_i)>u} = 1 - u. \hfill\Box$
%%
%%%\subsection{Multivariate Conditional Survival Function\label{subsec:condsurv}}
%%Equation \ref{eqn:chi_ij} implies that $\chi_{\jmath\ell}>0$, and so, no asymptotic independence is possible under our proposed model. For the analysis of extreme values it is of interest to calculate the multivariate conditional survival function. The following result provides the relevant expression, as a function of the angular measure.
%%\begin{prop}
%%Assume the same conditions of Proposition \ref{ppchi}. 
%%Let $\alpha \subset \{1, \ldots ,d\}$ be a collections of indexes. 
%%Then     
%%\begin{equation} \label{eqn:condsurv}
%%    \begin{aligned}
%%    &\hspace{-0.5cm}\mathrm{Pr}\left[\bigcap_{\ell \in \alpha} Z_\ell > z_\ell \mid \bigcap_{\ell\not\in\alpha} Z_\ell > z_\ell\right] \\
%%    &= \frac{\expect{\bigwedge_{k = 1}^d 1\wedge \frac{V_k}{z_k}}}{\expect{\bigwedge_{k \not\in\alpha}1\wedge\frac{V_k}{z_k}}}.
%%    \end{aligned}
%%  \end{equation}
%%\end{prop}  
%%The proof uses a similar approach to the proof of Proposition \ref{ppchi}.
%%
  % EOF

>>>>

During the last decade or so, much work has been done in the exploration of the 
    definition and properties of an appropriate generalization of the univariate 
    GP distribution to a multivariate setting.  To mention some of the papers on 
    the topic, the work of \citep{rootzen2006} defines the generalized Pareto 
    distribution, with further analysis on these classes of distributions 
    presented in \cite{falk2008} and \cite{michel2008}.  A recent review of the 
    state of the art in multivariate peaks over threshold modelling using 
    generalized Pareto is provided in \cite{rootzen2018} while \cite{RoSeWa2018a} 
    provides insight on the theoretical properties of possible parametrizations. 
    These are use in \cite{KiRoSeWa2019} for likelihood-based models for PoT 
    estimation. A frequently used method for describing the dependence in 
    multivariate distributions is to use a copula. \cite{renard2007}, and 
    \cite{falk2019} provide successful  examples of this approach in an EVT 
    framework. \cite{ferreira2014} presents a constructive definition of the 
    Pareto process, that generalizes the GP to an infinite dimensional setting. It 
    consists of decomposing the process into independent radial and angular 
    components. Such an approach can be used in the finite dimensional case, where 
    the angular component contains the information pertaining to the dependence 
    structure of the random vector. Based on this definition, we present a novel 
    approach for modelling  the angular component with families of distributions 
    that provide flexibility and can be applied in a moderately large dimensional 
    setting.  Our focus on the angular measure is similar to that in \cite{boldi2007},
    \cite{SaNa2014} and \cite{HaCaCh2017}, that consider Bayesian non-parametric 
    approaches. Yet, our approach differs in that it is established in the 
    peaks-over-threshold regime, and avoids the problem of dealing with the so 
    called moment constraint by considering a constructive definition of the 
    multivariate GP, based on the infinity norm.  Further, the

@@

\section{A multivariate PoT model\label{sec:multivariatepot}}
To develop a multivariate PoT model for extreme values consider a 
    $d$-dimensional random vector $\bm{W} = (W_1, \ldots ,W_d)$ with
    cumulative distribution $F$. Following \cite{RoSeWa2018a}, assume 
    that there exists sequences of vectors $\bm{a}_n$ and $\bm{b}_n$,
    and a $d$-variate distribution $G$ such that 
    $\lim_{n\rightarrow\infty} F^n(\bm{a}_n \bm{w} + \bm{b}_n) = 
    G(\bm{w})$. $G$ is a $d$-variate generalized extreme value 
    distribution. It follows that
    \begin{equation*}
        \begin{aligned}
        \lim\limits_{n\rightarrow\infty} &\prob{\bm{a}^{-1}_n (\bm{W} - \bm{b}_n) 
      \leq \bm{w} \mid \bm{W} \not\leq \bm{b}_n)} \\ 
        &\hspace{0.5cm}= \frac{\log G(\bm{w} \wedge \bm{0}) 
        - \log G(\bm{w}) }{\log G(\bm{0})} = H(\bm{w}).
        \end{aligned}
    \end{equation*}
    $H$ a multivariate Pareto distribution, and corresponds to the joint
    distribution conditional on exceeding a multivariate threshold. In addition to
    a copula function, $H$ depends on two $d$-dimensional vectors of 
    parameters that we denote as $\bm{\xi}$ for the shapes and $\bm{\sigma}$ 
    for the scales. \cite{RoSeWa2018a} provide a number of stochastic 
    representations for $H$.  In particular, denote as $Z$ a
    random variable with distribution $H$ where $\bm{\xi}= 1$ and 
    $\bm{\sigma} = 0$.  Then, Remark 1 justifies the representation in 
    \cite{ferreira2014}, giving $\bm{Z} = \bm{V}R$
    where $R$ and $\bm{V}$ are independent. $R = \|\bm{Z}\|_\infty$ is 
    distributed as a standard Pareto random variable, and $\bm{V} = 
    \bm{Z}/\|\bm{Z}\|_\infty$ is a random vector in  
    $\mathbb{S}_{\infty}^{d-1}$, the positive orthant of the 
    unit sphere under $\mathcal{L}_{\infty}$ norm, with distribution $\Phi$. 
    This representation is central to the methods proposed in this paper.
    $R$ and $\bm{V}$ are referred to, respectively, as the {\em radial} 
    and {\em angular} components of $H$. The angular measure controls 
    the dependence  structure of $\bm{Z}$  in the tails. In view of 
    this, to obtain a  PoT model we seek a flexible model for the 
    distribution of $\bm{V} \in {\mathbb S}_{\infty}^{d-1}$.

@@

% EOF
